Preliminary Draft

### 2021-07-26 article/draft.tex
some updates in the discussion

In this article we have assessed three broad categories of interventions to tackle misinformation by Facebook, Twitter and YouTube: suspension of accounts, introduction of notices and flags and reducing the visibility of content. We have provided a methodology that can be useful for third-party monitoring. The code needed to collect the data and create the figures is available on the following public GitHub repository: \href{https://github.com/medialab/truth-and-trust-online-2021/}{https://github.com/medialab/truth-and-trust-online-2021/}, and we hope it can be useful to those working on addressing online misinformation. In the light of the present research, we discuss in this section data and policy related recommandations.

@@

Mainstream platforms such as Facebook, Twitter and Youtube are increasingly working towards higher data transparency (see links that redirect to the respective transparency centers in Appendix \ref{links}). Yet the present research sheds light on the lack and the splintering of data, which impede the academic community from studying successfully online misinformation and assessing the impact of platforms' interventions regarding this phenomenon. Hence, few comments are in order regarding the lack of data, the lack of transparency regarding misinformation related interventions and the splintering of data.

@@

{\bf Lack of data.} First, when Facebook, Twitter or YouTube intervene by suspending an account, a page or a channel, the data disappears. It can no longer be accessed via scraping or via the official APIs (Twitter, YouTube, CrowdTangle) for the purposes of scientific research. This hampers the investigation of how platforms moderate content. To further illustrate this point, in a related research project (see Thero and Vincent (2020)~\cite{therovincent}) we searched on CrowdTangle for public accounts sharing specific content associated with misinformation in November $2020$. We selected $94$ Facebook pages corresponding to our criteria. We then tried to collect the posts on these pages in January $2021$. But we discovered that $11$ pages could no longer be found and that the related data had disappeared from the CrowdTangle API. Similarly, when a YouTube channel gets suspended or a YouTube video is deleted the data related to them is removed from the YouTube API. For instance, the original list of YouTube videos from table \ref{tab2} with failed fact-checks, had over $200$ YouTube videos available in March $2021$. However, by June $2021$, thirty videos had disappeared both from YouTube and the YouTube API for policy violation.

@@

Aside from the lack of data due to suspensions, some essential metrics for the study of misinformation are not accessible via main APIs. In particular, the $reach$ of posts on Facebook {\color{red}(i.e. the number of people who actually see a given post)} cannot be recovered from the CrowdTangle API and researchers can only build proxies for this metric from the available data (see section \ref{reduce_fb} for an example). Furthermore, to the best of our knowledge, fields related to flags, notices and information panels are not available on the official APIs (with the only exceptions of two types of interstitials in Twitter API). For the purposes of the present research we thus had to scrape this information from Facebook, Twitter and YouTube (see section \ref{flags}). Again, this disrupts the monitoring of platforms' interventions concerning misinformation and the assessment of their impact {\color{red} Misinformation review REF + others to show that we are not the only ones to say that}.

{\bf Lack of transparency.} Second, mainstream platforms generally lack transparency regarding their misinformation related interventions.  The broad lines of policies regarding misinformation are communicated via blog posts and are not typically part of the set of rules or community guidelines of mainstream platforms (see Appendix \ref{links}). {\color{red} For example, Facebook's strategy to tackle ``False News'' is three fold : Remove, Reduce, Inform. It is explained via a blog post on the facebook Newsroom. Similarly, Twitter communicates about actions related to misinformation via their \href{https://blog.twitter.com/en\_us/authors.TwitterSafety}{Twitter Safety} Blog/account. In a post on YouTube Official Blog, the platform explains its ``Four Rs of Responsibility" and how it raises authoritative content, reduces borderline content and harmful misinformation (see table \ref{summary} in Appendix \ref{links}). Nevertheless, when navigating through the categories of policy areas, to the best of our knowledge, misinformation is absent from the data available on the transparency centers of Facebook, Twitter and YouTube.} To provide an example, the $2021$ $Q1$ {\it Community Standards Enforcement Report} of Facebook, provides the volume of actioned content only regarding the following policy areas: (1) Adult Nudity \& Sexual Activity, (2) Bullying \& Harassement, (3) Child Nudity \& Sexual Exploitation, (4) Dangerous Orgs: Organized Hate, (5) Dangerous Orgs: Terrorism, (6) Fake Accounts, (7) Hate Speech, (8) Regulated Good: Drugs, (9) Regulated Good: Firearms, Spam, (10) Suicide and Self-injury, (11) Violent \& Graphic Content. More precisely, the academic community can only be aware of interventions related to misinformation via monitoring social media accounts related to domain names with several failed fact-checks or via articles in News outlets. For example, while conducting the present research, it was brought to our attention that during our study, the Twitter account related to the domain name $globalresearch.ca$ was suspended. Yet, there was no official communication or blog post by Twitter Safety team regarding this intervention (see section \ref{globalresearch}).

We are aware that full transparency regarding operational aspects of misinformation related interventions can backfire. This is because, the prospective aimed users might find means to circumvent the partly automated interventions. {\color{red} To provide an example, it is very likely that YouTube is allocating an information panel on all YouTube videos mentioning COVID-19 and related terms in their title (see section \ref{youtube_panels}). Yet, we noticed during the present study, that some YouTube channels are trying to avoid the likely automated information panels, by using notational variants such as ``C.O.V.I.D'' or ``C O V I D''. Similarly, we came across  notational variants on Twitter of domain names with several failed fact-checks. To provide an example, Twitter users introduce  the following notational variants when sharing an article from the website\footnote{Here is an example of a failed fact-check of one of the articles of the website {\it off-guardian.org} : \href{https://www.politifact.com/factchecks/2020/jul/07/blog-posting/covid-19-tests-are-not-scientifically-meaningless/}{politifact.com/factchecks/2020/jul/07/blog-posting/covid-19-tests-are-not-scientifically-meaningless/}} {\it off-guardian.org}: {\it off-guardian. org}, {\it off-guardian .org}, {\it off-guardian./org}, {\it off-guardian*org}. For the case of this domain name, we successfully managed to post a Tweet containing the domain name {\it off-guardian.org}, which we deleted after a few seconds. In other words, users can include in their Tweets links containing  {\it off-guardian.org} and only a warning message appears when a user clicks on the link.} For this case, we failed to understand precisely what kind of intervention the users were circumventing. Hence, a balance is needed between: $(i)$ fully transparent communication about misinformation related interventions so that the users can understand how the platforms are regulating their content, and $(ii)$ not explicitly giving too much information which would allow the policies to be bypassed.

@@

{\bf Scattered data.} Third, to study online misinformation, researchers, NGOs and data journalists turn to multiple APIs besides the official ones managed by the mainstream platforms themselves (e.g. BuzzSumo). They do so, because multiple data sources can be complementary and no specific API has complete data. {\color{red} vs. scraping. [to be drafted by HT... and cite the article ~\cite{api} about APIs]...., article du NYTimes sur CT, APIs qui ouvrent et qui ferment, limites de l'API de Twitter entre la V1 et la V2.}

@@

{\bf Lack of coordination between platforms.}

{\bf Policies recommandations.}

\smallskip
%\subsection{Policy recommendations}

In this section, we further provide some policy recommendations. It is very likely that interventions used to tackle misinformation ...

%what do i want to say ?

@@

\noindent * Brouillon / esquisse / organisation des bullet points: (N.B. the bullet point ${\color{green}\surd}$ means that the item/idea was used in the above paragraphs)

@@

\item[Idea 1:] Not an exhaustive list of interventions + brief summary of what we do.

	\item[${\color{green}\surd}$] We do not provide  an exhaustive list of methods on how to investigate the platforms’ policies, as other APIs and some databases could be useful for that matter. We rather provide a methodology to investigate key policies, that can be useful to researchers or journalists interested in implementing external monitoring. For a detailed list of interventions of multiple platforms we invite the reader to check the following \href{https://airtable.com/shrO0ooI9WSEfIUhb/tblAWQwFOiihKdQjm/viwZLAOzLK1NQ0c2n?blocks=hide}{airtable} compiled by Slatz and Leibowicz (2021)~\cite{niemanlab}.

	\item[${\color{green}\surd}$] We hope this paper can illustrate how the platforms policies can be verified. All the code needed to collect the data, and create the figures can be found on a public GitHub repo: \href{https://github.com/medialab/truth-and-trust-online-2021/}{https://github.com/medialab/truth-and-trust-online-2021/}, so that other researchers and journalists can easily implement and reproduce our methodology for the purposes of bringing to the society insights concerning the debate around misinformation.

\item[Idea 2:] Lack of data. Scattered data, indirect methods.

		\item[${\color{green}\surd}$] For a previous research project (add reference to the article) we searched on CrowdTangle for public accounts sharing specific content associated with misinformation in November $2020$, and selected $94$ Facebook pages corresponding to our criteria. We then tried to collect these pages' posts in January $2021$, and discovered that $11$ pages could not be found anymore. This highlights an important issue when studying misinformation trends on Facebook: some data disappears from the CrowdTangle API as accounts are deleted or changed to {\it private}.

		\item[${\color{green}\surd}$] After a channel gets suspended or a video is deleted the data related to them is removed from the YouTube API. This is a problem that affected the study to investigate the suspension reasons or the removal. For instance, the original list of videos from table 1 had more than 200 videos available in March 2021, however, by June 2021 30 videos got deleted from YouTube because they contain content that is against the YouTube guidelines and their data got removed from the API.

		\item[${\color{green}\surd}$] Pas accès au ``reach" !  ranking not available. Fb nous a donné certaines données (les données Condor) et dedans on a le nombre de clicks par exemple

		\item[${\color{green}\surd}$] Researchers should be able to access the data of deleted accounts on the main platforms.

		\item[${\color{green}\surd}$] Currently the flag or notice banners data is not present in the platforms’ API datasource, and this prevents researchers and journalists from easily investigating that matter. They have to use or build a scraper to access the data.

		\item[${\color{green}\surd}$] Lack of transparency in the platforms’ communication that makes their policies harder to verify.

		\item[${\color{green}\surd}$] To facilitate the verification in the policy applications, we would generally recommend for the platforms to be more transparent. But too much transparency on how the regulation policies are exactly implemented can actually backfire. For example YouTube is certainly applying an 'information' banner on all videos mentioning COVID-19 and related terms in their title. Misinformation accounts are trying to avoid the official banners by using terms as 'C.O.V.I.D' or 'C O V I D'. If YouTube was totally transparent on that matter and published the list of 'dangerous' words that leads to an information banner, this list would of course help us to understand YouTube's policies but it would also help the misinformation actors to escape the regulation. There is thus a balance between communicating enough so the public can know precisely how the platforms are regulating their content, but without giving too much information that would allow the policies to be bypassed.

		\item[${\color{green}\surd}$] Make point about the communication of platforms for their policy: 4R of YouTube, RRI of Facebook. Facebook's strategy to tackle ``False News'' is three fold : Remove, Reduce, Inform. It is explained via a blog post on the facebook Newsroom. Similarly, Twitter communicates about actions related to misinformation via their \href{https://blog.twitter.com/en\_us/authors.TwitterSafety}{Twitter Safety} Blog/account. In a post on YouTube Official Blog, the platform explained its ``Four Rs of Responsibility" and how it raises authoritative content, reduces borderline content and harmful misinformation (see table \ref{summary} in Appendix \ref{links}).

		\item There are other ways to collect data from platforms, and besides Buzzsumo, other API are also aggregating data from multiple social platforms. For example Newsguard, blablabla... In this tweet, we can see the interface of XXX being used in this study from a data journalist: \href{https://twitter.com/Shayan86/status/1394742784683298818/photo/2}{ex}

		\item Finally, we found irregularities in the number of Infowars articles collected from Buzzsumo. While Infowars usually publishes $20$ to $30$ articles per day, only $53$ articles were collected in the $31$-day period between June $11$ to July $11$, 2020. A temporary crawling problem coming from Buzzsumo may have caused this lack of data. Because no database is perfect, we would like to highlight the importance of cross-checking information between different sources when possible.

>>>>

In this article we have assessed three broad categories of interventions to tackle misinformation by Facebook, Twitter and YouTube: suspension of accounts, introduction of notices and flags and reducing the visibility of content. We have provided a methodology that can be useful for third-party monitoring. The code needed to collect the data and create the figures is available on the following public GitHub repository: \href{https://github.com/medialab/truth-and-trust-online-2021/}{https://github.com/medialab/truth-and-trust-online-2021/}, and we hope it can be a useful ressource for those working on addressing online misinformation.

@@

In the light of the present research, we discuss in this section data and policy related recommandations. Mainstream platforms such as Facebook, Twitter and Youtube are increasingly working towards higher data transparency (see links that redirect to the respective transparency centers in Appendix \ref{links}). Yet the present research sheds light on the lack and the splintering of data, which impede the academic community from studying successfully online misinformation and assessing the impact of platforms' interventions regarding this phenomenon. Hence, few comments are in order regarding the lack of data, the lack of transparency regarding misinformation related interventions, the splintering of data and the lack of interventions designed specifically for misinformation.

@@

{\bf Lack of data.} First, when Facebook, Twitter or YouTube intervene by suspending an account, a page or a channel, the data disappears. It can no longer be accessed via scraping or via the official APIs (Twitter, YouTube, CrowdTangle) for the purposes of scientific research. This hampers the investigation of how platforms moderate content. To further illustrate this point, the original list of YouTube videos summarized in table \ref{tab2} with failed fact-checks, had over $200$ YouTube videos available in March $2021$. However, by June $2021$, thirty videos had disappeared both from YouTube and the YouTube API for policy violation. Similarly, in a related research project (see Thero and Vincent (2020)~\cite{therovincent}) we searched on CrowdTangle for $94$ public accounts sharing specific content associated with misinformation in November $2020$. We tried to collect the posts from those $94$ pages in January $2021$, to only discover that $11$ pages had disappeared from the CrowdTangle API.

@@

Aside from the lack of data due to suspensions, some essential metrics for the study of misinformation are not accessible via main APIs. In particular, the $reach$ of posts on Facebook, which is the number of people who actually see a given post, cannot be recovered from the CrowdTangle API. Hence, researchers can only build proxies for this metric based on available data (see section \ref{reduce_fb} for an example). Furthermore, to the best of our knowledge, fields related to flags, notices and information panels are not available on the official APIs; with the exception of the {\it withheld}  and {\it possibly sensitive content} interstitials in the Twitter API v2. For the purposes of the present research we thus had to scrape this information from Facebook, Twitter and YouTube. Again, this disrupts the monitoring of platforms' interventions concerning misinformation and the assessment of their impact. We refer the reader to the insightful commentary by Pasquetto, Swire-Thompson, Amazeen et al. (2020)~\cite{pasquetto} and the references therein, for a discussion on partial or biased social media data.

%{\color{red} Misinformation review REF + others to show that we are not the only ones to say that}.
\smallskip

{\bf Lack of transparency.} Second, mainstream platforms generally lack transparency regarding their misinformation related interventions.  The broad lines of policies regarding misinformation are communicated via blog posts and are not typically part of the set of rules or community guidelines of mainstream platforms (see Appendix \ref{links}). Yet, when navigating through the categories of policy areas, to the best of our knowledge, misinformation is absent from the data available on the transparency centers of Facebook, Twitter and YouTube. To illustrate, the $2021$ $Q1$ {\it Community Standards Enforcement Report} of Facebook, provides the volume of actioned content only regarding the following policy areas: (1) Adult Nudity \& Sexual Activity, (2) Bullying \& Harassement, (3) Child Nudity \& Sexual Exploitation, (4) Dangerous Orgs: Organized Hate, (5) Dangerous Orgs: Terrorism, (6) Fake Accounts, (7) Hate Speech, (8) Regulated Good: Drugs, (9) Regulated Good: Firearms, Spam, (10) Suicide and Self-injury, (11) Violent \& Graphic Content. More precisely, platforms' official communication about misinformation interventions is scarce and the academic community, NGOs and data journalists, usually discover interventions related to misinformation via monitoring social media accounts related to domain names with several failed fact-checks or via articles in News outlets. For example, while conducting the present research, it was brought to our attention that during our study, the Twitter account related to the domain name $globalresearch.ca$ was suspended. But, there was no official communication or blog post by Twitter Safety team regarding this intervention (see section \ref{globalresearch}).

We are aware that full transparency regarding operational aspects of misinformation related interventions can backfire. This is because, the prospective aimed users might find means to circumvent the partly automated interventions. We noticed that some YouTube channels are trying to avoid the likely automated COVID-19 information panels, by using notational variants such as ``C.O.V.I.D'' or ``C O V I D''. Similarly, we came across notational variants on Twitter of the domain name {\it off-guardian.org}\footnote{Here is an example of a failed fact-check of one of the articles of the website {\it off-guardian.org} : \href{https://www.politifact.com/factchecks/2020/jul/07/blog-posting/covid-19-tests-are-not-scientifically-meaningless/}{politifact.com/factchecks/2020/jul/07/blog-posting/covid-19-tests-are-not-scientifically-meaningless/}}, such as {\it off-guardian. org}, {\it off-guardian .org}, {\it off-guardian./org}, {\it off-guardian*org}.\footnote{For this case, we failed to understand precisely what kind of intervention Twitter users were circumventing, since they could share in a Tweet an article redirecting to the website {\it off-guardian.org} with the correct notation and only a warning message would appear when one clicks on it.} Hence, a balance is needed between: $(i)$ fully transparent communication about misinformation related interventions so that the users can understand how the platforms are regulating their content, and $(ii)$ not explicitly giving too much information which would allow the policies to be bypassed.

@@

{\bf Scattered data.} Third, to study online misinformation, researchers, NGOs and data journalists turn to multiple APIs besides the official ones managed by the mainstream platforms themselves (e.g. BuzzSumo). They do so, because multiple data sources can be complementary and no specific API has complete data. {\color{pink} To be drafted by HT:  vs. scraping, cite the article ~\cite{api} about APIs, article du NYTimes sur CT, APIs qui ouvrent et qui ferment, limites de l'API de Twitter entre la V1 et la V2.}

@@

{\bf Lack of misinformation specific interventions.}

@@

%\noindent * Brouillon / esquisse / organisation des bullet points: (N.B. the bullet point ${\color{green}\surd}$ means that the item/idea was used in the above paragraphs)

@@

%\item[Idea 1:] Not an exhaustive list of interventions + brief summary of what we do. 
%	
%	\begin{itemize}
%
%	\item[${\color{green}\surd}$] We do not provide  an exhaustive list of methods on how to investigate the platforms’ policies, as other APIs and some databases could be useful for that matter. We rather provide a methodology to investigate key policies, that can be useful to researchers or journalists interested in implementing external monitoring. For a detailed list of interventions of multiple platforms we invite the reader to check the following \href{https://airtable.com/shrO0ooI9WSEfIUhb/tblAWQwFOiihKdQjm/viwZLAOzLK1NQ0c2n?blocks=hide}{airtable} compiled by Slatz and Leibowicz (2021)~\cite{niemanlab}.
%
%	\item[${\color{green}\surd}$] We hope this paper can illustrate how the platforms policies can be verified. All the code needed to collect the data, and create the figures can be found on a public GitHub repo: \href{https://github.com/medialab/truth-and-trust-online-2021/}{https://github.com/medialab/truth-and-trust-online-2021/}, so that other researchers and journalists can easily implement and reproduce our methodology for the purposes of bringing to the society insights concerning the debate around misinformation. 
%	
%	\end{itemize}

%\item[Idea 2:] Lack of data. Scattered data, indirect methods.  
%	
%	
%	\begin{itemize}

%		\item[${\color{green}\surd}$] For a previous research project (add reference to the article) we searched on CrowdTangle for public accounts sharing specific content associated with misinformation in November $2020$, and selected $94$ Facebook pages corresponding to our criteria. We then tried to collect these pages' posts in January $2021$, and discovered that $11$ pages could not be found anymore. This highlights an important issue when studying misinformation trends on Facebook: some data disappears from the CrowdTangle API as accounts are deleted or changed to {\it private}. 
%		\item[${\color{green}\surd}$] After a channel gets suspended or a video is deleted the data related to them is removed from the YouTube API. This is a problem that affected the study to investigate the suspension reasons or the removal. For instance, the original list of videos from table 1 had more than 200 videos available in March 2021, however, by June 2021 30 videos got deleted from YouTube because they contain content that is against the YouTube guidelines and their data got removed from the API. 
%		\item[${\color{green}\surd}$] Pas accès au ``reach" !  ranking not available. Fb nous a donné certaines données (les données Condor) et dedans on a le nombre de clicks par exemple
%		\item[${\color{green}\surd}$] Researchers should be able to access the data of deleted accounts on the main platforms.
%		\item[${\color{green}\surd}$] Currently the flag or notice banners data is not present in the platforms’ API datasource, and this prevents researchers and journalists from easily investigating that matter. They have to use or build a scraper to access the data.
%		\item[${\color{green}\surd}$] Lack of transparency in the platforms’ communication that makes their policies harder to verify.
%		\item[${\color{green}\surd}$] To facilitate the verification in the policy applications, we would generally recommend for the platforms to be more transparent. But too much transparency on how the regulation policies are exactly implemented can actually backfire. For example YouTube is certainly applying an 'information' banner on all videos mentioning COVID-19 and related terms in their title. Misinformation accounts are trying to avoid the official banners by using terms as 'C.O.V.I.D' or 'C O V I D'. If YouTube was totally transparent on that matter and published the list of 'dangerous' words that leads to an information banner, this list would of course help us to understand YouTube's policies but it would also help the misinformation actors to escape the regulation. There is thus a balance between communicating enough so the public can know precisely how the platforms are regulating their content, but without giving too much information that would allow the policies to be bypassed.
%		\item[${\color{green}\surd}$] Make point about the communication of platforms for their policy: 4R of YouTube, RRI of Facebook. Facebook's strategy to tackle ``False News'' is three fold : Remove, Reduce, Inform. It is explained via a blog post on the facebook Newsroom. Similarly, Twitter communicates about actions related to misinformation via their \href{https://blog.twitter.com/en\_us/authors.TwitterSafety}{Twitter Safety} Blog/account. In a post on YouTube Official Blog, the platform explained its ``Four Rs of Responsibility" and how it raises authoritative content, reduces borderline content and harmful misinformation (see table \ref{summary} in Appendix \ref{links}). 
%		\item There are other ways to collect data from platforms, and besides Buzzsumo, other API are also aggregating data from multiple social platforms. For example Newsguard, blablabla... In this tweet, we can see the interface of XXX being used in this study from a data journalist: \href{https://twitter.com/Shayan86/status/1394742784683298818/photo/2}{ex}
%		\item Finally, we found irregularities in the number of Infowars articles collected from Buzzsumo. While Infowars usually publishes $20$ to $30$ articles per day, only $53$ articles were collected in the $31$-day period between June $11$ to July $11$, 2020. A temporary crawling problem coming from Buzzsumo may have caused this lack of data. Because no database is perfect, we would like to highlight the importance of cross-checking information between different sources when possible.
%	\end{itemize}
